The Hidden Manifold Distance for Functional Data

### 2019-08-30 paperAAAI.tex
final writing touches

% The file aaai20.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
% command to correct it. You may not alter the value below 2.5 in
\title{The Hidden Manifold Distance for Functional ata}
%Your title must be in mixed case, not sentence case. 
% That means all verbs (including short verbs like be, is, using,and go), 
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\author{Susan Wei\textsuperscript{\rm 1} and Marie-H\'el\`{e}ne Descary\textsuperscript{\rm 2}\\ % All authors must be in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line
\textsuperscript{\rm 1} The University of Melbourne, \textsuperscript{\rm 2}Universit\'e du Qu\'ebec \`{a} Montr\'eal\\ %If you have multiple authors and multiple affiliations
% use superscripts in text and roman font to identify them. For example, Sunil Issar,\textsuperscript{\rm 2} J. Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}. Note that the comma should be placed BEFORE the superscript for optimum readability
 % email address must be in roman text type, not monospace or sans serif
}
\begin{document}

@@

\begin{abstract}
In many machine learning applications, the observed data is a
discretely-sampled realization of a continuous process. While one can
naively treat such data as vectors in Euclidean space, one may do better
to regard them as functional data. For certain types of functional data,
there is further structure to exploit, e.g.~probability density
functions or warped curves of a common template function. Namely, the
underlying curves of interest may lie on a manifold of low intrinsic
dimension. This work addresses the estimation of pairwise geodesic
distances between functional manifold data that are observed with noise,
causing them to lie off the manifold. This setting falls outside of
classic manifold learning techniques which require data to live exactly
on or very near a manifold. The proposed methodology first sends the
observed functional data to the hidden manifold, estimated using
subspace-constrained mean-shift. Geodesic distances are subsequently
calculated by employing shortest-path algorithms on this estimated
manifold. Improved estimation of the pairwise geodesic distance has
beneficial implications for many downstream tasks in functional data
analysis which we illustrate in the case of distance-based functional
classification/clustering.
\end{abstract}

@@

>>>>

% The file aaai20.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
% command to correct it. You may not alter the value below 2.5 in
\title{The Hidden Manifold Distance for Functional Data}
%Your title must be in mixed case, not sentence case. 
% That means all verbs (including short verbs like be, is, using,and go), 
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\author{Susan Wei\textsuperscript{\rm 1} and Marie-H\'el\`{e}ne Descary\textsuperscript{\rm 2}\\ % All authors must be in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line
\textsuperscript{\rm 1} The University of Melbourne, \textsuperscript{\rm 2}Universit\'e du Qu\'ebec \`{a} Montr\'eal\\ %If you have multiple authors and multiple affiliations
% use superscripts in text and roman font to identify them. For example, Sunil Issar,\textsuperscript{\rm 2} J. Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}. Note that the comma should be placed BEFORE the superscript for optimum readability
 % email address must be in roman text type, not monospace or sans serif
}
\begin{document}

@@

\begin{abstract}
In many machine learning applications, the observed data is a
discretely-sampled realisation of a continuous process. While one can
naively treat such data as vectors in Euclidean space, one may do better
to regard them as functional data.  
When the underlying curves of interest lie on a manifold of low intrinsic
dimension, e.g.~probability density
functions or warped curves of a common template function, there is further structure to exploit. 
This work addresses the estimation of pairwise geodesic
distances between functional manifold data that are observed with noise,
causing them to lie off the manifold. This setting falls outside of
classic manifold learning techniques which require data to live exactly
on or very near a manifold. The proposed methodology first sends the
observed functional data to the hidden manifold, estimated using
subspace-constrained mean-shift. Geodesic distances are subsequently
calculated by employing shortest-path algorithms on this estimated
manifold. Improved estimation of the pairwise geodesic distance has
potential benefits for downstream tasks such as distance-based functional
???Marie: update classification/clustering???.
\end{abstract}

@@

\input{MarieDataAnalysis}

\section{Acknowledgement}\label{acknowledgement}

The authors would like to acknowledge support from the ARC Centre of
Excellence for Mathematical and Statistical Frontiers (ACEMS) which
enabled the two authors to work on the project at the University of
Melbourne.

\section{Appendices}\label{appendices}

The result of the square root velocity manifold scenario is presented in
Figure \ref{fig:betas}. The Spline L2 method has similar performance to
our proposed method. This could be due to the fact that we did not
sample the manifold broadly enough so that the sample lives in a locally
flat region where the \(L^2\) distance holds.

\begin{figure}
\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_MSE}
\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_isometry}
\includegraphics[height=0.3\textheight]{sim_results/sce=4_SNR=high_Kobs=30_Pearson}
\caption{Manifold of square root Beta densities. Besides Spline IsoGeo, all methods perform comparably well.}
\label{fig:betas}
\end{figure}

\bibliography{references}
\bibliographystyle{aaai}
